Freezing of a Weeks-Chandler-Anderson Fluid using Classical Density Functional Theory

### 2020-12-13 papers/thesis-kirstie/thesis.tex
pres edits
went back to previous intro with more detail on FMT

\documentclass[double,12pt]{beavtex}
\usepackage{graphicx}
\usepackage{rotating} %Package added to allow the rotation of figures and chart on a page, {sidewaysfigure} command
\usepackage{tablefootnote} %Packaged added to allow footnotes in the tabular environment, use \tablefootnote command
\title{Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory}
\author{Kirstie L. Finster}
\degree{Master of Science}
\doctype{Thesis}
\department{Physics}
\depttype{Department}
\depthead{Head}
\major{Physics}
\advisor{David Roundy}
\submitdate{December 10, 2020}
\commencementyear{2021}

@@

The goal of this project is to generate temperature-density 
and pressure-temperature phase diagrams for a Weeks-Chandler-Anderson 
(WCA) model fluid using classical Density Functional Theory (cDFT). 
To do this, the Helmholtz free energies of the fluid
in a liquid state and in a crystalline solid state must be found.
In general, the Helmholtz free energy, which consists of an ideal part 
and an excess part, can be written as a functional of the number density 
profile $n(\vec{r})$ which describes the way the average number 
density varies spatially. 
\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}

Soft Fundamental Measure Theory (SFMT) will be used to form the 
functional for the excess Helmholtz free energy. 
SFMT is a method that makes the task of 
forming an expression for the excess Helmholtz free energy of a particular 
model fluid much easier since the interaction that each sphere, or atom, 
has with every other sphere does not have to be taken explicitly into account.
Once the functional for the Helmholtz free energy is formed, then 
according to cDFT, the equilibrium
Helmholtz free energy of the system can be found for a particular 
temperature and density by varying the number 
density profile $n(\vec{r})$ until the free energy is minimized. 
Phase diagrams can then be constructed using Maxwell's Double Tangent Construction. 
Results will be compared to those from Monte-Carlo simulations.

>>>>

\documentclass[double,12pt]{beavtex}
\usepackage{graphicx}
\usepackage{rotating} %Package added to allow the rotation of figures and chart on a page, {sidewaysfigure} command
\usepackage{tablefootnote} %Packaged added to allow footnotes in the tabular environment, use \tablefootnote command
\title{Freezing of a Weeks-Chandler-Anderson Fluid using Classical Density Functional Theory}
\author{Kirstie L. Finster}
\degree{Master of Science}
\doctype{Thesis}
\department{Physics}
\depttype{Department}
\depthead{Head}
\major{Physics}
\advisor{David Roundy}
\submitdate{December 10, 2020}
\commencementyear{2021}

@@

Classically, fluids have been modeled by a collection of small, 
impenetrable spheres colliding with one another like billiard balls. 
In the simplest case, the spheres are treated as having only a replulsive 
potential. This model, called the hard-sphere fluid, has served as the 
primary reference system for fluids. This is due to the fact that the 
repulsion between atoms is much stronger than their attraction, and so 
the repulsive potential, which is descriptive of hard spheres, dominates 
the interactions between atoms. In 1954 Zwanzig introduced the Square-well 
fluid which is a perturbation of the hard-sphere fluid created by adding 
a hard-edge attractive potential to the repulsive potential of the 
hard-sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
spheres with softly fading replusive potentials which he modeled by making
the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
This reflects the idea that soft-spheres penetrate more deeply into 
each other when they collide at higher temperatures.

In 1967 Barker and Henderson combined the perturabtion approach of Zwanzig 
with Rowlinson's idea of using temperature-dependent radii to model spheres 
with soft repulsive potentials to create the Barker-Henderson model 
fluid~\cite{barker1967perturbation}. In this model, a soft-sphere fluid is 
modeled by using a hard-sphere fluid with hard spheres of larger radii 
at low temperatures, and smaller radii at high temperatures. In 1971 Weeks, 
Chandler, and Anderson used temperature and density dependent hard-sphere 
radii to model spheres with the more realistic softly fading repulsive 
portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
which is known as the Weeks-Chandler-Anderson model fluid.

ADD MORE HERE TO COME UP TO DATE!!

While models for particular fluids were being created, powerful methods 
to describe the properties of fluids were also being developed. 
In the late 1950s, Scaled Particle Theory (SPT)~\cite{ReissSPT} was 
introduced from which an equation of state, and an expression for the 
excess Helmholtz free energy of a homogeneous hard-Sphere fluid can be 
obtained~\cite{Hansen}. 
%\textcolor{red}{
In 1958, Percus-Yevick introduced an alternative method 
for the hard-sphere fluid based on an approximate closure to the 
Ornstein-Zernike integral equation~\cite{PercusYevick}.
%Percus and Yevick took another approach to the hard-Sphere fluid.
In 1961 Morita and Hiroike introduced a density functional 
approach~\cite{MoritaDFT} to finding the equilibrium Grand and Helmholtz free energies, 
and the cooresponding equilibrium number density profile for 
inhomogeneous fluids. To use this classical Density Functional Theory (cDFT) 
approach, an expression for the free energy 
as a functional of an average number density that varies spatially 
must be constructed for a particular fluid.
%DFT Evans 1979
%Led by the consistency bewteen results from SPT and Percus-Yevick, 
In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
(which can be constructed from SPT~\cite{rosenfeld1989,Hansen,santos2012phi3}) 
to form a functional for the hard-sphere fluid.
FMT is based on the idea that for a hard-sphere fluid the
lowest term in the virial expansion of the excess chemical potential can 
be expressed in terms of a sum of weighted densities using a weight 
function of the form of a backward Heavyside step function (shown below) 
that can be expressed in terms of a sum of convolutions involving 
a set of weight functions that can be related to the geometry of a sphere.
\begin{displaymath}
   \theta(|\vec r - \vec r'|-(\vec R_i + \vec R_j)) = \end{displaymath}
\begin{equation} w_i^{(3)} x w_j^{(0)} + w_j^{(3)} x w_i^{(0)} + w_i^{(1)} x w_j^{(2)} + w_j^{(1)} x w_i^{(2)} + \vec w_i^{(1)} x \vec w_j^{(2)} + \vec w_j^{(1)} x \vec w_i^{(2)}
\end{equation}
Tarazona later improved on the 
method by introducing a tensor version of one of the energy density terms 
in the functional~\cite{tarazonaphi3, 
santos2012phi3}, and in 1999 Schmidt adopted FMT for spheres with soft 
interatomic potential energies
(which rise to infinity over a distance, rather than instantaneously 
like the potential energy between hard spheres)
forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}.

%by replacing $w^{(3)}$ with a weight function that would result in  
%soft interatomic potentials (which rise to infinity over a distance, 
%rather than instantaneously).

Schmidt applied his Soft Fundamental Measure Theory to various fluids 
with soft repulsive and soft attractive potentials, and developed an error function  
model 
%based on a Gaussian weight function 
for spheres with steep, short-range soft repulsive 
potentials~\cite{schmidt2000fluid}.
%, similar to that exhibited by a WCA fluid. 
In this paper, clasical Density Functional 
Theory will be used to predict the freezing of a Weeks-Chandler-Anderson 
model fluid by applying SFMT with Schmidt's proposed error function model, 
in which we have incorporated our temperature dependent parameters to fit the 
functional to the soft-sphere WCA fluid.
%form the functional for the excess Helmholtz free energy. 
%for generating the weight funcions, and Tarazona's tensor functional. 
Results will be compared to those generated from Monte-Carlo simulations.

%Results will be compared to those from Monte-Carlo simulations.

%Newer Intro....
%The goal of this project is to generate temperature-density 
%and pressure-temperature phase diagrams for a Weeks-Chandler-Anderson 
%(WCA) model fluid using classical Density Functional Theory (cDFT). 
%To do this, the Helmholtz free energies of the fluid
%in a liquid state and in a crystalline solid state must be found.
%In general, the Helmholtz free energy, which consists of an ideal part 
%and an excess part, can be written as a functional of the number density 
%profile $n(\vec{r})$ which describes the way the average number 
%density varies spatially. 
%\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}

%Soft Fundamental Measure Theory (SFMT) will be used to form the 
%functional for the excess Helmholtz free energy. 
%SFMT is a method that makes the task of 
%forming an expression for the excess Helmholtz free energy of a particular 
%model fluid much easier since the interaction that each sphere, or atom, 
%has with every other sphere does not have to be taken explicitly into account.
%Once the functional for the Helmholtz free energy is formed, then 
%according to cDFT, the equilibrium
%Helmholtz free energy of the system can be found for a particular 
%temperature and density by varying the number 
%density profile $n(\vec{r})$ until the free energy is minimized. 
%Phase diagrams can then be constructed using Maxwell's Double Tangent Construction. 
%Results will be compared to those from Monte-Carlo simulations.
